The Weakly Nonlinear Magnetorotational Instability

### 2016-08-29 paper/widegap_mri_paper.tex
insulating adjoint bc's added to appendix
test hmri
ncc test w jeff's params

and $B = 0$ at $r = r_1, r_2$. Here, $I_n$, $K_n$ are the modified Bessel functions of the first and second kind, respectively.

@@

When $\xi$ in Equation \ref{eq:unperturbed_matrix_eqns} is not equal to zero, the helical MRI arises. We examine a single fiducial helical MRI case, for the parameters listed in Table \ref{table:parameters}. At the conclusion of the weakly nonlinear analysis, we find that the coefficients of Equation \ref{eq:gle} are $b = $, $d = $, and $c = $, i.e. complex. The marginal helical MRI is thus described by a complex GLE. This difference in character between the amplitude equations that modulate the weakly nonlinear standard and helical MRI is a consequence of the same property that makes the helical MRI an overstability.

@@

\beq
\langle \mathbf{V^\dagger} | (\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} \rangle = \langle (\mathcal{L}^\dagger + \mathcal{H}^\dagger) \mathbf{V}^\dagger | \mathbf{V} \rangle,
\eeq

@@

\beq\label{eq:inner_product_def}
\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \frac{k_c}{2\pi} \int_{-\pi/k_c}^{\pi/k_c} \int_{r_1}^{r_2} \mathbf{V}^{\dagger*} \cdot \mathcal{L} \mathbf{V} \, \mathrm{d} r \mathrm{d} z
\eeq

We find that

\beq
\mathcal{L}^\dagger = \mathcal{L}_0^\dagger - \partial_z \mathcal{L}_1^\mathrm{T} + d_z^2 \mathcal{L}_2^\mathrm{T} - \partial_z^3 \mathcal{L}_3^\mathrm{T} + \partial_z^4 L_4^\mathrm{T},
\eeq

@@

\beq
\mathcal{H}^\dagger = - d_z \mathcal{H}^\mathrm{T}
\eeq

@@

\beq
\mathcal{L}_0^\dagger = \left[\begin{matrix}
-\frac{1}{\reye} (-\frac{3}{r^4} \partial_r + \frac{3}{r^3}\partial_r^2 - \frac{2}{r^2}\partial_r^3 + \frac{1}{r}\partial_r^4) & 0 & 0 & 0 \\
0 & -\frac{1}{\reym} (\partial_r^2 - \frac{1}{r} \partial_r) & 0 & 0 \\
0 & 0 &-\frac{1}{\reye} (\partial_r^2 + \frac{1}{r}\partial_r - \frac{1}{r^2}) & 0 \\
0 & 0 & 0 & -\frac{1}{\reym} (\partial_r^2 - \frac{1}{r} \partial_r) \\
\end{matrix}\right].
\eeq

@@

\beq
(\mathcal{L}^\dagger + \xi \widetilde{\mathcal{H}}^\dagger)\mathbf{V}^\dagger = 0
\eeq

@@

Note that the condition on $B^\dagger$ differs from that on $B$. We take the inner product of the adjoint homogenous solution with Equation \ref{eq:ordere3} to find

@@

>>>>

and $B = 0$ at $r = r_1, r_2$ (see \citei{Willis:2002bh}). Here, $I_n$, $K_n$ are the modified Bessel functions of the first and second kind, respectively.

@@

When $\xi$ in Equation \ref{eq:unperturbed_matrix_eqns} is not equal to zero, the helical MRI arises. We examine a single fiducial helical MRI case, for the parameters listed in Table \ref{table:parameters}. At the conclusion of the weakly nonlinear analysis, we find that the coefficients of Equation \ref{eq:gle} are $b = $, $d = $, and $c = $, i.e. complex. The marginal helical MRI is thus described by a complex GLE. This difference in character between the amplitude equations that modulate the weakly nonlinear standard and helical MRI is a consequence of the same property that makes the helical MRI an overstability. With the introduction of an azimuthal component, the background magnetic field acquires a handedness that is not present in a purely axial field.

%************* APPENDICES ************************%

@@

\beq
\langle \mathbf{V^\dagger} | (\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} \rangle = \langle (\mathcal{L}^\dagger + \xi \mathcal{H}^\dagger) \mathbf{V}^\dagger | \mathbf{V} \rangle,
\eeq

@@

\beq\label{eq:inner_product_def}
\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \frac{k_c}{2\pi} \int_{-\pi/k_c}^{\pi/k_c} \int_{r_1}^{r_2} \mathbf{V}^{\dagger*} \cdot \mathcal{L} \mathbf{V} \, r \mathrm{d} r \mathrm{d} z
\eeq

We derive the adjoint operator by successive integration by parts, to find

\beq
\mathcal{L}^\dagger = \mathcal{L}_0^\dagger - \partial_z \mathcal{L}_1^\dagger + d_z^2 \mathcal{L}_2^\dagger - \partial_z^3 \mathcal{L}_3^\dagger + \partial_z^4 L_4^\dagger
\eeq

@@

\beq
\mathcal{H}^\dagger = - d_z \mathcal{H}^\mathrm{T},
\eeq

@@

\beq
\mathcal{L}_0^\dagger = \left[\begin{matrix}
-\frac{1}{\reye} (-\frac{3}{r^5} + \frac{3}{r^4}\partial_r - \frac{3}{r^3} \partial_r^2 + \frac{2}{r^2}\partial_r^3 + \frac{1}{r}\partial_r^4 ) & 0 & 0 & 0 \\
0 & -\frac{1}{\reym} (\partial_r^2 - \frac{1}{r} \partial_r) & 0 & 0 \\
0 & 0 &-\frac{1}{\reye} (\partial_r^2 + \frac{1}{r}\partial_r - \frac{1}{r^2}) & 0 \\
0 & 0 & 0 & -\frac{1}{\reym} (\partial_r^2 - \frac{1}{r} \partial_r) \\
\end{matrix}\right].
\eeq

@@

\beq\label{eq:adjoint}
(\mathcal{L}^\dagger + \xi \widetilde{\mathcal{H}}^\dagger)\mathbf{V}^\dagger = 0
\eeq

@@

Note that the condition on $B^\dagger$ differs from that on $B$. For the insulating boundary conditions we apply to the helical MRI, the magnetic boundary conditions applied to Equation \ref{eq:adjoint} are

\beq
B^\dagger = (- k \frac{I_0 (k r)}{I_1 (k r)} + \frac{1}{r} + \partial_r ) A^\dagger = 0 \, \, \mathrm{at} \, \, r = r_1
\eeq

\beq
B^\dagger = (k \frac{K_0 (k r)}{K_1 (k r)} + \frac{1}{r} + \partial_r ) A^\dagger = 0 \, \, \mathrm{at} \, \, r = r_2
\eeq

We take the inner product of the adjoint homogenous solution with Equation \ref{eq:ordere3} to find

@@

\bibitem[{Willis \& Barenghi(2002)}]{Willis:2002bh}
Willis, A.P. \& Barenghi, C.F. 2002, JFM 463, 361
